Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-04-08 VegOzone.tex
Updated abstract

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and air pollution profiles on the modeled susceptibility of subarctic vegetation to ozone. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of near-surface ozone, e.g. due to wildfires.
We have monitored near-surface ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available near-surface ozone products for the region. Over all, ozone concentrations did not differ significantly between the two years. While temperatures and global irradiance diverged substantially from the multi annual mean precipitation varied only to some extend. 
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial.
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatologically reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased in 2018; corroborated by evidence of visible injury in the Svanhovd ozone garden.
\end{abstract}

>>>>

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of near-surface ozone, e.g. due to wildfires.\\
We have monitored near-surface ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on probability density function (PDF) and with respect to the Swedish and Finnish atmospheric monitoring sites. We review available near-surface ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years, but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above ground biomass reduction in perannial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}
